% Requires:
% \usepackage{amsmath,amssymb,amsthm,bm}

\section{Two-level configuration-conditional spectral calibration model}

\subsection{Purpose and scope}

The goal is to learn a calibration system that is trained offline from
\emph{colocated calibration campaigns}, in which all participating sensors
observe the same latent spectrum, and is then deployed online to calibrate
\emph{individual sensors} operating at arbitrary locations, where different
sensors generally observe different spectra.

The model therefore has two levels:

\begin{enumerate}
    \item an \emph{offline identification level}, which uses same-scene
    multi-sensor campaigns to learn persistent sensor- and
    configuration-dependent calibration laws;
    \item an \emph{online deployment level}, which applies the learned
    persistent laws to calibrate a single sensor in an arbitrary scene,
    without requiring any common-field assumption at deployment time.
\end{enumerate}

The key structural principle is that only the offline level uses the
``all sensors see the same spectrum'' assumption. The online level does not.

\subsection{Offline calibration corpus}

Let
\[
\mathcal D_{\mathrm{cal}}
=
\{\mathcal C_m\}_{m=1}^{M_{\mathrm{cal}}}
\]
be the corpus of offline calibration campaigns.
Campaign $m$ is represented by
\[
\mathcal C_m
=
\bigl(
\mathcal S_m,\,
K_m,\,
f_m,\,
c_m,\,
Y_m
\bigr),
\]
where:

\begin{itemize}
    \item $\mathcal S_m \subseteq \{1,\dots,S_{\max}\}$ is the set of
    sensors participating in campaign $m$;
    \item $M_m := |\mathcal S_m| \ge 2$ is the number of participating sensors;
    \item $K_m \in \mathbb N$ is the number of aligned acquisitions;
    \item
    \[
    f_m=(f_{m,1},\dots,f_{m,J_m}),
    \qquad
    f_{m,1}<\cdots<f_{m,J_m},
    \]
    is the retained common frequency grid for campaign $m$;
    \item
    \[
    c_m \in \mathcal X
    \]
    is the campaign acquisition configuration, assumed constant within the
    campaign;
    \item
    \[
    c_m
    =
    \bigl(
    f_{{\mathrm c},m},\,
    B_m,\,
    \mathrm{RBW}_m,\,
    g_{{\mathrm{LNA}},m},\,
    g_{{\mathrm{VGA}},m}
    \bigr),
    \]
    where $f_{{\mathrm c},m}$ is central frequency, $B_m$ is span/sample-rate
    metadata, $\mathrm{RBW}_m$ is the resolution bandwidth,
    $g_{{\mathrm{LNA}},m}$ is the LNA gain, and
    $g_{{\mathrm{VGA}},m}$ is the VGA gain;
    \item
    \[
    Y_m=\{Y_{m,s,k,j}\}
    \]
    is the aligned PSD tensor in linear power units, indexed by
    \[
    s\in\mathcal S_m,\qquad
    k\in\{1,\dots,K_m\},\qquad
    j\in\{1,\dots,J_m\}.
    \]
\end{itemize}

\subsection{Offline same-scene observation model}

For campaign $m$, sensor $s\in\mathcal S_m$, acquisition $k$, and frequency
bin $j$, the observed PSD is modeled as
\begin{equation}
Y_{m,s,k,j}
=
G_{m,s,j}\,S_{m,k,j}
+
N_{m,s,j}
+
\varepsilon_{m,s,k,j},
\label{eq:offline-forward}
\end{equation}
where:

\begin{itemize}
    \item $S_{m,k,j}\ge 0$ is the latent common spectrum of the colocated
    calibration campaign at acquisition $k$ and frequency bin $j$;
    \item $G_{m,s,j}>0$ is the multiplicative gain of sensor $s$ in campaign
    $m$ at bin $j$;
    \item $N_{m,s,j}\ge 0$ is the additive floor of sensor $s$ in campaign
    $m$ at bin $j$;
    \item $\varepsilon_{m,s,k,j}$ is a residual term.
\end{itemize}

Conditionally on the latent spectrum and calibration parameters, the working
likelihood is Gaussian:
\begin{equation}
\varepsilon_{m,s,k,j}
\mid
S_{m,k,j},G_{m,s,j},N_{m,s,j}
\sim
\mathcal N\!\left(0,\sigma^2_{m,s,j}\right),
\label{eq:offline-gaussian}
\end{equation}
with $\sigma^2_{m,s,j}>0$.

\subsection{Persistent laws and campaign-specific deviations}

The two-level model separates calibration into:

\begin{enumerate}
    \item \emph{persistent sensor/configuration laws}, which are intended to
    transfer to deployment;
    \item \emph{campaign-specific deviations}, which account for residual
    nuisance variation during offline training.
\end{enumerate}

\subsubsection{Persistent latent laws}

Let
\[
G_s^{(0)}(f,c) > 0
\]
be a nominal multiplicative baseline. In the simplest current convention one
may set
\[
G_s^{(0)}(f,c)\equiv 1.
\]

Define three persistent latent functions:
\[
\Gamma_s^\star(f,c)\in\mathbb R,
\qquad
H_s^\star(f,c)\in\mathbb R,
\qquad
V_s^\star(f,c)\in\mathbb R,
\]
where:

\begin{itemize}
    \item $\Gamma_s^\star(f,c)$ is the persistent log-gain law;
    \item $H_s^\star(f,c)$ is the persistent additive-floor pre-activation;
    \item $V_s^\star(f,c)$ is the persistent variance pre-activation.
\end{itemize}

The persistent deployment-scale laws are then
\begin{align}
G_s^\star(f,c)
&=
G_s^{(0)}(f,c)\exp\!\bigl(\Gamma_s^\star(f,c)\bigr),
\label{eq:persistent-gain}
\\
N_s^\star(f,c)
&=
\phi\!\bigl(H_s^\star(f,c)\bigr),
\label{eq:persistent-floor}
\\
\sigma_s^{2,\star}(f,c)
&=
\sigma_{\min}^2+\phi\!\bigl(V_s^\star(f,c)\bigr),
\label{eq:persistent-var}
\end{align}
where
\[
\phi(x)=\log(1+e^x)
\]
is the softplus map and $\sigma_{\min}^2>0$ is a fixed minimum variance.

\subsubsection{Neural parameterization of the persistent laws}

Let $e_s\in\mathbb R^{d_s}$ be a trainable embedding for sensor identity
$s$, and let
\[
u_\theta:\mathcal X\to\mathbb R^{d_c}
\]
be a trainable configuration encoder.

Let
\[
b^\Gamma_\ell:\mathbb R\to\mathbb R,
\quad \ell=1,\dots,L_\Gamma,
\]
\[
b^H_\ell:\mathbb R\to\mathbb R,
\quad \ell=1,\dots,L_H,
\]
\[
b^V_\ell:\mathbb R\to\mathbb R,
\quad \ell=1,\dots,L_V,
\]
be fixed smooth frequency basis functions (for example, cubic B-splines).

The persistent laws are parameterized by a neural network with global weights
$\theta$:
\begin{align}
a^\Gamma_\theta(s,c)
&=
A^\Gamma_\theta\!\bigl(e_s,u_\theta(c)\bigr)
\in\mathbb R^{L_\Gamma},
\label{eq:aGamma}
\\
a^H_\theta(s,c)
&=
A^H_\theta\!\bigl(e_s,u_\theta(c)\bigr)
\in\mathbb R^{L_H},
\label{eq:aH}
\\
a^V_\theta(s,c)
&=
A^V_\theta\!\bigl(e_s,u_\theta(c)\bigr)
\in\mathbb R^{L_V},
\label{eq:aV}
\end{align}
and
\begin{align}
\Gamma_s^\star(f,c)
&=
\sum_{\ell=1}^{L_\Gamma}
a^\Gamma_{\theta,\ell}(s,c)\,b^\Gamma_\ell(f),
\label{eq:Gamma-star-basis}
\\
H_s^\star(f,c)
&=
\sum_{\ell=1}^{L_H}
a^H_{\theta,\ell}(s,c)\,b^H_\ell(f),
\label{eq:H-star-basis}
\\
V_s^\star(f,c)
&=
\sum_{\ell=1}^{L_V}
a^V_{\theta,\ell}(s,c)\,b^V_\ell(f).
\label{eq:V-star-basis}
\end{align}

Thus one global neural system parameterized by $\theta$ generates smooth
sensor- and configuration-conditional calibration laws over frequency.

\subsubsection{Campaign-specific deviations}

For offline campaign $m$, define campaign-specific residual corrections
\[
\delta^\Gamma_{m,s,j}\in\mathbb R,
\qquad
\delta^H_{m,s,j}\in\mathbb R,
\qquad
\delta^V_{m,s,j}\in\mathbb R.
\]

Then the campaign-specific parameters used in
\eqref{eq:offline-forward}--\eqref{eq:offline-gaussian} are
\begin{align}
G_{m,s,j}
&=
G_s^{(0)}(f_{m,j},c_m)\,
\exp\!\Bigl(
\Gamma_s^\star(f_{m,j},c_m)
+
\delta^\Gamma_{m,s,j}
\Bigr),
\label{eq:campaign-gain}
\\
N_{m,s,j}
&=
\phi\!\Bigl(
H_s^\star(f_{m,j},c_m)
+
\delta^H_{m,s,j}
\Bigr),
\label{eq:campaign-floor}
\\
\sigma^2_{m,s,j}
&=
\sigma_{\min}^2
+
\phi\!\Bigl(
V_s^\star(f_{m,j},c_m)
+
\delta^V_{m,s,j}
\Bigr).
\label{eq:campaign-var}
\end{align}

The role of $\delta^\Gamma,\delta^H,\delta^V$ is to absorb campaign-specific
nuisance variation such as local thermal conditions, residual hardware state,
or other nonpersistent effects during offline identification.

\subsection{Identifiability}

\subsubsection{Local campaign-scale nonidentifiability}

For a fixed campaign $m$, the observation model is invariant under the
transformation
\[
G_{m,s,j}\mapsto a_{m,j}G_{m,s,j},
\qquad
S_{m,k,j}\mapsto a_{m,j}^{-1}S_{m,k,j},
\qquad
a_{m,j}>0.
\]
Hence the multiplicative scale is not identifiable without an explicit
reference convention.

\subsubsection{Global deployment-scale convention}

To define a persistent deployment scale, choose positive weights
\[
\pi_s>0,
\qquad
\sum_{s=1}^{S_{\max}}\pi_s=1,
\]
and impose the global reference convention
\begin{equation}
\sum_{s=1}^{S_{\max}}
\pi_s\,\Gamma_s^\star(f,c)=0
\qquad
\text{for all }(f,c).
\label{eq:global-reference-convention}
\end{equation}

Equivalently,
\[
\prod_{s=1}^{S_{\max}}
\left(
G_s^\star(f,c)
\right)^{\pi_s}
=
\prod_{s=1}^{S_{\max}}
\left(
G_s^{(0)}(f,c)
\right)^{\pi_s}.
\]

If $G_s^{(0)}(f,c)\equiv 1$, then
\begin{equation}
\prod_{s=1}^{S_{\max}}
\left(
G_s^\star(f,c)
\right)^{\pi_s}
=
1.
\label{eq:geom-mean-one-global}
\end{equation}

This fixes the persistent calibration laws on a global network reference
scale. In implementation, \eqref{eq:global-reference-convention} can be
enforced by subtracting the weighted mean across sensors from the raw neural
log-gain outputs.

\subsubsection{Mean-zero campaign deviations}

To prevent the campaign residual corrections from reintroducing arbitrary
multiplicative scale, impose
\begin{equation}
\frac{1}{M_m}
\sum_{s\in\mathcal S_m}
\delta^\Gamma_{m,s,j}
=
0,
\qquad
j=1,\dots,J_m,
\quad
m=1,\dots,M_{\mathrm{cal}}.
\label{eq:campaign-mean-zero-delta}
\end{equation}

Under \eqref{eq:campaign-mean-zero-delta}, the latent common spectrum
$S_{m,k,j}$ of each calibration campaign is expressed on the persistent
global deployment scale defined by
\eqref{eq:global-reference-convention}.

\paragraph{Remark.}
If an absolute metrological reference is available, then
\eqref{eq:global-reference-convention} may be replaced by an absolute anchor,
for example by fixing
\[
\Gamma_{s_{\mathrm{ref}}}^\star(f,c)=0
\]
for a trusted reference instrument or by adding supervised absolute-loss
terms. Without such information, the framework is rigorous only on the chosen
reference scale.

\subsection{Offline latent-spectrum update}

For fixed values of $(\theta,\delta^\Gamma,\delta^H,\delta^V)$, the latent
common spectrum in calibration campaign $m$ is estimated by weighted least
squares.

Define
\[
\omega_{m,s,j}
=
\frac{1}{\sigma_{m,s,j}^2}.
\]
Then
\begin{equation}
\widehat S_{m,k,j}
=
\left[
\frac{
\sum_{s\in\mathcal S_m}
\omega_{m,s,j}\,G_{m,s,j}\,
\bigl(
Y_{m,s,k,j}-N_{m,s,j}
\bigr)
}{
\sum_{s\in\mathcal S_m}
\omega_{m,s,j}\,G_{m,s,j}^2
}
\right]_+,
\label{eq:latent-update}
\end{equation}
where $[x]_+=\max\{x,0\}$ or, in implementation, clipping to a small positive
numerical floor $\varepsilon_S>0$.

\subsection{Offline objective}

Let
\[
\Delta
=
\bigl\{
\delta^\Gamma_{m,s,j},
\delta^H_{m,s,j},
\delta^V_{m,s,j}
\bigr\}_{m,s,j}
\]
collect all campaign-specific deviations.

\subsubsection{Negative log-likelihood term}

The Gaussian negative log-likelihood is
\begin{equation}
\mathcal L_{\mathrm{nll}}(\theta,\Delta,S)
=
\frac{1}{2}
\sum_{m=1}^{M_{\mathrm{cal}}}
\sum_{s\in\mathcal S_m}
\sum_{k=1}^{K_m}
\sum_{j=1}^{J_m}
\left[
\log \sigma_{m,s,j}^2
+
\frac{
\bigl(
Y_{m,s,k,j}
-
G_{m,s,j}S_{m,k,j}
-
N_{m,s,j}
\bigr)^2
}{
\sigma_{m,s,j}^2
}
\right].
\label{eq:nll}
\end{equation}

\subsubsection{Smoothness of campaign deviations}

To regularize campaign-specific corrections across frequency, define a
second-difference operator $D_{m,2}$ on the frequency grid $f_m$.
For a uniform grid it may be the usual second difference; for a nonuniform
grid one should use the corresponding nonuniform second-difference operator.

The campaign-deviation regularizer is
\begin{equation}
\mathcal R_{\mathrm{dev}}(\Delta)
=
\sum_{m=1}^{M_{\mathrm{cal}}}
\sum_{s\in\mathcal S_m}
\Bigl[
\lambda^\Gamma_{\mathrm{smooth}}
\|D_{m,2}\delta^\Gamma_{m,s,\cdot}\|_2^2
+
\lambda^H_{\mathrm{smooth}}
\|D_{m,2}\delta^H_{m,s,\cdot}\|_2^2
+
\lambda^V_{\mathrm{smooth}}
\|D_{m,2}\delta^V_{m,s,\cdot}\|_2^2
\Bigr],
\label{eq:Rdev-smooth}
\end{equation}
where, for example,
\[
\delta^\Gamma_{m,s,\cdot}
=
(\delta^\Gamma_{m,s,1},\dots,\delta^\Gamma_{m,s,J_m})^\top.
\]

\subsubsection{Shrinkage of campaign deviations}

Because the campaign deviations are nuisance terms, they should shrink toward
zero:
\begin{equation}
\mathcal R_{\mathrm{shrink}}(\Delta)
=
\sum_{m=1}^{M_{\mathrm{cal}}}
\sum_{s\in\mathcal S_m}
\Bigl[
\lambda^\Gamma_0
\|\delta^\Gamma_{m,s,\cdot}\|_2^2
+
\lambda^H_0
\|\delta^H_{m,s,\cdot}\|_2^2
+
\lambda^V_0
\|\delta^V_{m,s,\cdot}\|_2^2
\Bigr].
\label{eq:Rdev-shrink}
\end{equation}

\subsubsection{Weight decay on the persistent model}

The persistent neural model is regularized by
\begin{equation}
\mathcal R_\theta(\theta)
=
\lambda_\theta \|\theta\|_2^2.
\label{eq:Rtheta}
\end{equation}

\subsubsection{Optional reliable-sensor anchor}

If campaign $m$ provides a reliable sensor index $s_m^\star\in\mathcal S_m$,
one may add a weak anchor on its campaign residual gain correction:
\begin{equation}
\mathcal R_{\mathrm{rel}}(\Delta)
=
\sum_{m=1}^{M_{\mathrm{cal}}}
\lambda_{\mathrm{rel}}
\sum_{j=1}^{J_m}
\bigl(
\delta^\Gamma_{m,s_m^\star,j}
\bigr)^2.
\label{eq:Rrel}
\end{equation}
This is optional and should remain soft.

\subsubsection{Total offline training objective}

The baseline offline objective is
\begin{equation}
\mathcal J_{\mathrm{off}}(\theta,\Delta,S)
=
\mathcal L_{\mathrm{nll}}(\theta,\Delta,S)
+
\mathcal R_{\mathrm{dev}}(\Delta)
+
\mathcal R_{\mathrm{shrink}}(\Delta)
+
\mathcal R_\theta(\theta)
+
\mathcal R_{\mathrm{rel}}(\Delta),
\label{eq:Joff}
\end{equation}
subject to the identifiability constraints
\eqref{eq:global-reference-convention} and
\eqref{eq:campaign-mean-zero-delta}.

\subsection{Offline estimation problem}

Without any additional same-scene agreement penalty, the offline fit is the
constrained optimization problem
\begin{equation}
(\widehat\theta,\widehat\Delta,\widehat S)
\in
\operatorname*{arg\,min}_{\theta,\Delta,S}
\mathcal J_{\mathrm{off}}(\theta,\Delta,S)
\quad
\text{subject to }
\eqref{eq:global-reference-convention}
\text{ and }
\eqref{eq:campaign-mean-zero-delta}.
\label{eq:offline-estimation-problem}
\end{equation}

A practical implementation may use block alternation:

\begin{enumerate}
    \item for fixed $(\theta,\Delta)$, update $S$ by
    \eqref{eq:latent-update};
    \item for fixed $S$, update $(\theta,\Delta)$ by gradient-based
    optimization of \eqref{eq:Joff};
    \item project the persistent and campaign log-gain corrections onto the
    constraints \eqref{eq:global-reference-convention} and
    \eqref{eq:campaign-mean-zero-delta}.
\end{enumerate}

\subsection{Optional extension: same-scene consistency penalty}

The motivation is structural: each offline calibration campaign is a
\emph{same-scene} experiment, so all participating sensors observe the same
latent spectrum in \eqref{eq:offline-forward}. At deployment, however, only
the persistent laws transfer. It is therefore natural to regularize the
persistent calibration map itself so that, on the offline corpus, it produces
mutually consistent corrected spectra across sensors.

For campaign $m$, define the persistent deployment-scale laws evaluated on the
campaign grid:
\begin{align}
G_{m,s,j}^\star(\theta)
&=
G_s^{(0)}(f_{m,j},c_m)
\exp\!\bigl(\Gamma_{s,\theta}^\star(f_{m,j},c_m)\bigr),
\label{eq:consistency-persistent-gain}
\\
N_{m,s,j}^\star(\theta)
&=
\phi\!\bigl(H_{s,\theta}^\star(f_{m,j},c_m)\bigr).
\label{eq:consistency-persistent-floor}
\end{align}

Using only these persistent laws, define the persistent-only calibrated
spectrum
\begin{equation}
\widetilde S_{m,s,k,j}(\theta)
=
\frac{
\bigl[
Y_{m,s,k,j}
-
N_{m,s,j}^\star(\theta)
\bigr]_+
}{
G_{m,s,j}^\star(\theta)
}.
\label{eq:consistency-calibrated-spectrum}
\end{equation}

Under the same-scene assumption, an ideally calibrated model should satisfy
\[
\widetilde S_{m,s,k,j}(\theta)\approx S_{m,k,j}
\qquad
\text{for all }s\in\mathcal S_m,
\]
so the values
\(
\widetilde S_{m,s,k,j}(\theta)
\)
should agree across sensors for each aligned acquisition and frequency bin.

Because deployment diagnostics are naturally interpreted in a logarithmic
power scale, introduce a monotone transform
\begin{equation}
T:(0,\infty)\to\mathbb R,
\qquad
T(x)=\log(x+\varepsilon_T),
\qquad
\varepsilon_T>0.
\label{eq:consistency-transform}
\end{equation}
One may equivalently use $10\log_{10}(x+\varepsilon_T)$; this differs only by
a constant scaling that can be absorbed into the penalty weight.

Define
\begin{equation}
Z_{m,s,k,j}(\theta)
=
T\!\bigl(\widetilde S_{m,s,k,j}(\theta)\bigr),
\label{eq:consistency-log-spectrum}
\end{equation}
and let the cross-sensor consensus value be
\begin{equation}
\bar Z_{m,k,j}(\theta)
=
\frac{1}{M_m}
\sum_{s\in\mathcal S_m}
Z_{m,s,k,j}(\theta).
\label{eq:consistency-consensus}
\end{equation}

The same-scene consistency loss is then
\begin{equation}
\mathcal L_{\mathrm{cons}}(\theta)
=
\sum_{m=1}^{M_{\mathrm{cal}}}
\frac{1}{K_mJ_m}
\sum_{k=1}^{K_m}
\sum_{j=1}^{J_m}
\frac{1}{M_m}
\sum_{s\in\mathcal S_m}
\bigl(
Z_{m,s,k,j}(\theta)
-
\bar Z_{m,k,j}(\theta)
\bigr)^2.
\label{eq:Lcons}
\end{equation}
This penalty is zero if and only if, for every training campaign, aligned
acquisition, and frequency bin, the persistent-only corrected spectra
coincide across all participating sensors.

Moreover, for fixed $(m,k,j)$,
\begin{equation}
\sum_{\substack{s,s'\in\mathcal S_m\\ s<s'}}
\bigl(
Z_{m,s,k,j}
-
Z_{m,s',k,j}
\bigr)^2
=
M_m
\sum_{s\in\mathcal S_m}
\bigl(
Z_{m,s,k,j}
-
\bar Z_{m,k,j}
\bigr)^2.
\label{eq:consistency-pairwise-identity}
\end{equation}
Hence minimizing the consensus variance in \eqref{eq:Lcons} is equivalent,
up to a campaign-dependent constant factor, to minimizing the mean pairwise
squared discrepancy across sensors in the logarithmic power domain.

The augmented offline objective becomes
\begin{equation}
\mathcal J_{\mathrm{off}}^{\mathrm{cons}}(\theta,\Delta,S)
=
\mathcal J_{\mathrm{off}}(\theta,\Delta,S)
+
\lambda_{\mathrm{cons}}
\mathcal L_{\mathrm{cons}}(\theta),
\qquad
\lambda_{\mathrm{cons}}\ge 0,
\label{eq:Joff-cons}
\end{equation}
subject to the same identifiability constraints
\eqref{eq:global-reference-convention} and
\eqref{eq:campaign-mean-zero-delta}.

\paragraph{Interpretation.}
The role of $\mathcal L_{\mathrm{cons}}$ is not to replace the Gaussian
likelihood term, but to complement it with a criterion that directly matches
the physical structure of the offline corpus: after calibration, colocated
sensors should yield the same spectrum. The likelihood controls data fidelity
and noise weighting, whereas $\mathcal L_{\mathrm{cons}}$ explicitly promotes
\emph{post-calibration inter-sensor agreement} on the training campaigns.
This is precisely the quantity later inspected by before/after RMSE
diagnostics.

A crucial point is that the penalty is constructed from the \emph{persistent}
laws only, as in
\eqref{eq:consistency-persistent-gain}--\eqref{eq:consistency-calibrated-spectrum},
rather than from the full campaign-specific parameters. If the disagreement
penalty were imposed mainly through the nuisance deviations $\Delta$, those
terms could absorb the training disagreement without improving the calibration
map that is actually deployed. Thus
\eqref{eq:Joff-cons} encourages the part of the model that transfers online to
produce mutually consistent corrected spectra on same-scene offline campaigns.

\subsection{Persistent deployment laws after offline training}

Once the offline fit is complete, the deployed persistent laws are
\begin{align}
\widehat G_s^\star(f,c)
&=
G_s^{(0)}(f,c)
\exp\!\bigl(
\Gamma_{s,\widehat\theta}^\star(f,c)
\bigr),
\label{eq:deployed-persistent-gain}
\\
\widehat N_s^\star(f,c)
&=
\phi\!\bigl(
H_{s,\widehat\theta}^\star(f,c)
\bigr),
\label{eq:deployed-persistent-floor}
\\
\widehat \sigma_s^{2,\star}(f,c)
&=
\sigma_{\min}^2
+
\phi\!\bigl(
V_{s,\widehat\theta}^\star(f,c)
\bigr).
\label{eq:deployed-persistent-var}
\end{align}

These are the only quantities needed for online deployment in arbitrary
scenes.

\subsection{Online deployment on arbitrary scenes}

\subsubsection{Deployment data}

Consider a deployed sensor $s^\dagger\in\{1,\dots,S_{\max}\}$ operating under
configuration $c^\dagger$ on a retained frequency grid
\[
f^\dagger=(f^\dagger_1,\dots,f^\dagger_{J^\dagger}).
\]
Suppose it produces observations
\[
Y^\dagger=\{Y^\dagger_{k,j}\},
\qquad
k=1,\dots,K^\dagger,\quad
j=1,\dots,J^\dagger.
\]

No assumption is made that any other deployed sensor observes the same scene.

\subsubsection{Deployment observation model}

For deployment, the physically relevant model is
\begin{equation}
Y^\dagger_{k,j}
=
G^\dagger_j\,X^\dagger_{k,j}
+
N^\dagger_j
+
\varepsilon^\dagger_{k,j},
\label{eq:deployment-forward}
\end{equation}
where:
\begin{itemize}
    \item $X^\dagger_{k,j}$ is the \emph{sensor-local} latent spectrum of the
    deployed scene at that sensor;
    \item $G^\dagger_j$ and $N^\dagger_j$ are the persistent laws evaluated
    for that sensor and configuration;
    \item there is no common-spectrum relation tying this $X^\dagger$ to any
    other deployed sensor.
\end{itemize}

The deployed persistent calibration curves are
\begin{align}
\widehat G^\dagger_j
&=
\widehat G_{s^\dagger}^\star(f^\dagger_j,c^\dagger),
\label{eq:deployed-Gdag}
\\
\widehat N^\dagger_j
&=
\widehat N_{s^\dagger}^\star(f^\dagger_j,c^\dagger),
\label{eq:deployed-Ndag}
\\
\widehat \sigma^{2,\dagger}_j
&=
\widehat \sigma_{s^\dagger}^{2,\star}(f^\dagger_j,c^\dagger).
\label{eq:deployed-sigdag}
\end{align}

\subsubsection{Online calibration map}

The online calibrated PSD is
\begin{equation}
\widehat X^\dagger_{k,j}
=
\frac{
\bigl[
Y^\dagger_{k,j}
-
\widehat N^\dagger_j
\bigr]_+
}{
\widehat G^\dagger_j
}.
\label{eq:online-calibration-map}
\end{equation}

If one prefers unclipped subtraction, then replace $[\cdot]_+$ by identity and
handle negative corrected values downstream. However, nonnegativity clipping is
often numerically safer for PSD-like quantities.

\subsubsection{Per-bin deployment uncertainty}

A first-order variance propagation gives
\begin{equation}
\operatorname{Var}
\bigl(
\widehat X^\dagger_{k,j}
\mid
s^\dagger,c^\dagger
\bigr)
\approx
\frac{
\widehat \sigma^{2,\dagger}_j
}{
(\widehat G^\dagger_j)^2
}.
\label{eq:deployment-variance}
\end{equation}

This is the observation-noise contribution only. It does not include
epistemic uncertainty in $\widehat\theta$ unless such uncertainty is modeled
explicitly, for example through ensembling or variational methods.

\subsection{What transfers and what does not}

The same-scene assumption is used only in the offline calibration corpus to
identify persistent sensor/configuration behavior. It does \emph{not} appear in
the deployment rule \eqref{eq:online-calibration-map}.

Thus the transfer logic is:

\begin{enumerate}
    \item offline same-scene campaigns identify
    \[
    \widehat G_s^\star(f,c),\qquad
    \widehat N_s^\star(f,c),\qquad
    \widehat \sigma_s^{2,\star}(f,c);
    \]
    \item online deployment applies those laws to a single sensor in an
    arbitrary scene, with no requirement that other sensors observe the same
    latent spectrum.
\end{enumerate}

\subsection{Optional extension: slowly varying device-state covariates}

If slowly varying device-health covariates are available, such as temperature,
clock error diagnostics, age, or self-test features, let
\[
z_{m,s}\in\mathcal Z
\]
denote such covariates in offline campaign $m$ and
\[
z^\dagger\in\mathcal Z
\]
denote them at deployment time.

Then one may replace the persistent laws by
\[
\Gamma_s^\star(f,c,z),\qquad
H_s^\star(f,c,z),\qquad
V_s^\star(f,c,z),
\]
with the same structure as above. In that case the deployed laws become
\[
\widehat G_{s^\dagger}^\star(f^\dagger_j,c^\dagger,z^\dagger),\qquad
\widehat N_{s^\dagger}^\star(f^\dagger_j,c^\dagger,z^\dagger),
\]
and so on. This extension is optional; if such covariates are unavailable, one
simply omits $z$.

\subsection{Summary}

The two-level model consists of:

\begin{enumerate}
    \item a global persistent neural law
    \[
    (s,f,c)\mapsto
    \bigl(
    G_s^\star(f,c),
    N_s^\star(f,c),
    \sigma_s^{2,\star}(f,c)
    \bigr),
    \]
    learned offline from colocated same-scene campaigns;
    \item campaign-specific nuisance deviations used only during offline
    fitting;
    \item an online per-sensor calibration map
    \[
    \widehat X^\dagger_{k,j}
    =
    \frac{
    [Y^\dagger_{k,j}-\widehat N^\dagger_j]_+
    }{
    \widehat G^\dagger_j
    },
    \]
    which remains valid even when deployed sensors observe different spectra
    at different locations.
\end{enumerate}

Therefore the model is simultaneously:
\begin{itemize}
    \item \emph{configuration-conditional};
    \item \emph{sensor-specific};
    \item \emph{trained from same-scene campaigns};
    \item \emph{deployable to arbitrary scenes without a common-field assumption}.
\end{itemize}
